---
title: "CMVS Outcome Variables"
author: "Graham Albert"
date: "January 7, 2019"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

####First I will load the necessary packages in the work space.
####To load cvs files
```{r}
require(curl) 
```
######lto replace missing data
```{r}
require(mice)
```
######for advanced rotation methods
```{r}
require(GPArotation)
```
#####for data manipulation
```{r}
require(dplyr)
```
####for graphical depiction
```{r}
require(gplots)
require(gridExtra)
require(moments)
require(psych)##for EFA, and descriiptivves
require(BaylorEdPsych)####for missing value analysis
require(lavaan)####for latent variable analysis CFA and SEM.
require(semPlot)####for graphical depiction
require(BDgraph)##### for graphical depiction
require(yaml)
require(stringi)
```


####To load data set one for the CMVS.
```{r}
f <- curl("https://raw.githubusercontent.com/GrahamAlbert/Mate-Value/master/Components%20of%20self%20perceived%20mate%20value%20personality%20outcome%20variable%20and%20predictors_20190107.csv")
ICSSEM<- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
View(ICSSEM)
head(ICSSEM)
names(ICSSEM)
```

####To determine variable names of the data set.
```{r}
names(ICSSEM)[1]<-"ID"
names(ICSSEM)
```
####To obtain summary statistics for the data set.
```{r}
summary(ICSSEM)
```

####To obtain the descriptive statistics includinag the skewness and kurtosis values of the the cmvs
```{r}
require(psych)
normalityassessment = describe(ICSSEM[-c(1:5)])
View(normalityassessment)
```
####To generate histograms for the variables that are not normality distributed.
```{r}
hist(ICSSEM$Partner_Number)
hist(ICSSEM$Participant_Lifetime_Sex_Partners)
hist(ICSSEM$Past_Year_Sex_Partners)
hist(ICSSEM$Past_Month_Sex)
hist(ICSSEM$SOIR1)
```
####To mutate the data set to conduct a log 10 transformations for the skewed outcome variables
```{r}
ICSSEM2= mutate(ICSSEM, LGPartnerNumber=log10(Partner_Number+1), LGParticipantLifetimeSexPartners = log10(Participant_Lifetime_Sex_Partners+1),LGPastYearSexPartners=log10(Past_Year_Sex_Partners+1),LGPastMonthSex = log10(Past_Month_Sex+1),LGSOIR1 = log(SOIR1+1))
View(ICSSEM2)
ICSSEM2 = ICSSEM2[-c(5,6,9,10,11,65)]
View(ICSSEM2) 
```
####To generate histograms for the variables that are not normality distributed.
```{r}
hist(ICSSEM2$LGPartnerNumber)
hist(ICSSEM2$LGParticipantLifetimeSexPartners)
hist(ICSSEM2$LGPastYearSexPartners)
hist(ICSSEM2$LGPastMonthSex)
hist(ICSSEM2$LGSOIR1)
```

####To obtain descriptive stattistics including skew and kurtosis values of new data set including transformed varibales.
```{r}
normalityassessment2 = describe(ICSSEM2[-c(1:4)])
View(normalityassessment2)
names(ICSSEM2)
```
####To remove the CMVS from the data set.
```{r}
require(dplyr)
ICSSEM3=ICSSEM2[-c(5,7,34:59,68:91)] 
View(ICSSEM3)
```
####To obtain descriptive stattistics including skew and kurtosis values of new data set including transformed varibales.
```{r}
normalityassessment2 = describe(ICSSEM3[-c(1:4)])
View(normalityassessment2)
names(ICSSEM3)
require(BaylorEdPsych)
require(mvnmle)
LittleMCAR(ICSSEM3)
```


####To screening for missing values.
####To write a funtion to check for missng data.
```{r}
percentmiss=function(x){sum(is.na(x)/length(x))*100}
```


####To check for missing data by columns
```{r}
apply(ICSSEM3,2,percentmiss)
```
####It is less than 2% missing for all variables.

####To check for missing data by row
```{r}
missing=apply(ICSSEM3, 1,percentmiss)
table(missing)
```


####To exclude participants who are missing greater than 5% of their data.
```{r}
replacepeople=subset(ICSSEM3,missing<=5)
```
####I will be elimating three cases.

####Make sure that the columns are not missing too much data.
```{r}
apply(replacepeople, 2, percentmiss)
```

#####After excluding the cases with too much missing data we will replace the remaining missing data with a stocastic imputation method.
```{r}
require(mice)
tempnomiss = mice(replacepeople)
nomiss = complete(tempnomiss, 1)
summary(nomiss)
```
###To lookk at data within missing replaced.
```{r}
View(nomiss)
```
####To screen for multivariate outliers.
####To screen for multivariate outliers in the data set nomiss we will use mahalanobis distance statistic.
```{r}
mahal = mahalanobis(nomiss,
                    colMeans(nomiss),
                    cov(nomiss))
cutoff = qchisq(0.999, ncol(nomiss)) #####generates cutoff score
ncol(nomiss) #####determines df
table(mahal<cutoff)
```
####To exelucde the 17  outliers from the analysis.
```{r}
noout = subset(nomiss, mahal<cutoff)
View(noout)
```
####Therefore for this analysis we will have a total of 346 cases.


###Assumptions.
####To screen for multcorinately
####Althoughy the items are correlated (as they should) we do not have multi-colinearity so we proceed with SEM assumptionn set.
####I will be excluding the single report items from rows 1 to 4 from ID to whether participants are in a long term relationship from analysis.
```{r}
correl=cor(noout[-c(1:4)], use ="pairwise.complete.obs")
View(correl)
```
###EFA assumption set up.
```{r}
random = rchisq(nrow(noout[-c(1:4)]),7)
fake = lm(random ~.,data=noout[-c(1:4)])
standardized = rstudent(fake)
fitted = scale(fake$fitted.values)
```
####To test for normality.
```{r}
hist(standardized)
```
####The data is normal but slightly skewed to the right

####It appears as though we have multi-variate normality so we proceed to test linearity.
####To test for linearity.
```{r}
qqnorm(standardized)
abline(0,1)
```
####Our variables are linear so we procede to test homogenity of vairance.


####Homogenity of Variance.
```{r}
plot(fitted, standardized)
abline(0,0)
abline(v=0)
```




#####To create a general factor model > factor 1 and factor 2 specific.
####We will regress the general factor and the second factor onto the predictors.
```{r}
SEM.MODEL.BIFACTOR='
Generalfator=~ICS1+ICS2+ICS3+ICS5+ICS6+ICS7+ICS8+ICS12+ICS11+ICS10+ICS4+ICS9
factor1=~ICS1+ICS2+ICS3+ICS5+ICS6+ICS7+ICS8+ICS12+ICS11
factor2=~ICS10+ICS4+ICS9+ICS11
Generalfator~~0*factor1
Generalfator~~0*factor2
factor1~~0*factor2
Age_Sex~Generalfator+factor2+Participant_Sex
LGPartnerNumber~Generalfator+factor2+Participant_Age
LGParticipantLifetimeSexPartners~Generalfator+factor2+Participant_Age
LGPastYearSexPartners~Generalfator+factor2+Participant_LTR
LGPastMonthSex~Generalfator+factor2+Participant_LTR
Participant_LTR~Generalfator+factor2'
```
####To run the cfa for the ICS
```{r}
require(lavaan)
SEM.MODEL.BIFACTOR.FIT=sem(SEM.MODEL.BIFACTOR, data=noout,std.lv = TRUE, information="observed",orthogonal=TRUE)
```

####to depict the  models.
```{r}
require(BGgraphs)
require(semPlot)
semPaths(SEM.MODEL.BIFACTOR.FIT, whatLabels="std",layout="tree",edge.label.cex = 0.5,curvePivot = TRUE)
```

####To summarize model fit
```{r}
summary(SEM.MODEL.BIFACTOR.FIT, standardized=TRUE, rsquare=TRUE, fit.measure=TRUE)
```

###To produce standardized residuals for the
```{r}
zcorrel=residuals(SEM.MODEL.BIFACTOR.FIT,type='standardized')
View(zcorrel$cov)
```
####To obtain model indices of the SDQ.
```{r}
modindices(SEM.MODEL.BIFACTOR.FIT, sort. =TRUE, minimum.value=3.84)
```

####To obtian standarized and unstandardized factor loadings
```{r}
parameterEstimates(SEM.MODEL.BIFACTOR.FIT, standardized =T)
standardizedsolution(SEM.MODEL.BIFACTOR.FIT)
```



