---
title: "ICS CFA MGCFA"
author: "Graham Albert"
date: "September 10 2020"
output:pdf_document: default
---


####First I will load the necessary packages in the work space.
####To load cvs files
```{r}
require(curl) 
```
######lto replace missing data
```{r}
require(mice)
```
######for advanced rotation methods
```{r}
require(GPArotation)
```
#####for data manipulation
```{r}
require(dplyr)
```
####for graphical depiction
```{r}
require(gplots)
require(gridExtra)
require(moments)
require(psych)##for EFA, and descriiptivves
require(BaylorEdPsych)####for missing value analysis
require(lavaan)####for latent variable analysis CFA and SEM.
require(semPlot)####for graphical depiction
require(yaml)
require(stringi)
```


####To load data set one for the CMVS.
```{r}
f <- curl("https://raw.githubusercontent.com/GrahamAlbert/Mate-Value/master/Intrasexually_comeptitive_attitude_CFA_MGCFA_20190108_reduced.csv")
ICSCFA<- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
View(ICSCFA)
head(ICSCFA)
names(ICSCFA)
```

####To determine variable names of the data set.
```{r}
names(ICSCFA)[1]<-"ID"
names(ICSCFA)
```
####To obtain summary statistics for the data set.
```{r}
summary(ICSCFA)
```

####To obtain the descriptive statistics includinag the skewness and kurtosis values of the the ICS
```{r}
require(psych)
normalityassessment = describe(ICSCFA[-c(1:3)])
View(normalityassessment)
```
####All items are normally distributed.
####We will proceed to analyze patterns of missingness.

###After inspecting the skewness and kurtosis statistics for the cmvs it appears that all questions are relatively normally distributed therefore I will proceed with analyzing the data for missing values.
####To analysis the data set for missing value patters
```{r}
require(BaylorEdPsych)
require(mvnmle)
LittleMCAR(ICSCFA)
```





####To screening for missing values.
####To write a funtion to check for missng data.
```{r}
percentmiss=function(x){sum(is.na(x)/length(x))*100}
```


####To check for missing data by columns
```{r}
apply(ICSCFA,2,percentmiss)
```
####It is less than 2% missing for all variables.

####To check for missing data by row
```{r}
missing=apply(ICSCFA, 1,percentmiss)
table(missing)
```


####To exclude participants who are missing greater than 5% of their data.
```{r}
replacepeople=subset(ICSCFA,missing<=5)
```
####I will be elimating three cases.

####Make sure that the columns are not missing too much data.
```{r}
apply(replacepeople, 2, percentmiss)
```

#####After excluding the cases with too much missing data we will replace the remaining missing data with a stocastic imputation method.
```{r}
require(mice)
require(jomo)
set.seed(20)
tempnomiss = mice(replacepeople)
nomiss=complete(tempnomiss, 1)
summary(nomiss)
```
###To lookk at data within missing replaced.
```{r}
View(nomiss)
```
####To screen for multivariate outliers.
####To screen for multivariate outliers in the data set nomiss we will use mahalanobis distance statistic.
```{r}
mahal = mahalanobis(nomiss,
                    colMeans(nomiss),
                    cov(nomiss))
cutoff = qchisq(0.999, ncol(nomiss)) #####generates cutoff score
ncol(nomiss) #####determines df
table(mahal<cutoff)
```
####To exelucde the 14  outliers from the analysis.
```{r}
noout = subset(nomiss, mahal<cutoff)
View(noout)
```
####Therefore for this analysis we will have a total of 263 cases.


###Assumptions.
####To screen for multcorinately
####Althoughy the items are correlated (as they should) we do not have multi-colinearity so we proceed with CFA assumptionn set.
####I will be excluding the single report items from rows 1 and 2(sex and age) from analysis, so that I can look at the correlations amongst the items.
```{r}
correl=cor(noout[-c(1:3)], use ="pairwise.complete.obs")
View(correl)
```
###EFA assumption set up.
```{r}
random = rchisq(nrow(noout[-c(1:3)]),7)
fake = lm(random ~.,data=noout[-c(1:3)])
standardized = rstudent(fake)
fitted = scale(fake$fitted.values)
```
####To test for normality.
```{r}
hist(standardized)
```
####The data is normal but slightly skewed to the right

####It appears as though we have multi-variate normality so we proceed to test linearity.
####To test for linearity.
```{r}
qqnorm(standardized)
abline(0,1)
```
####Our variables are linear so we procede to test homogenity of vairance.


####Homogenity of Variance.
```{r}
plot(fitted, standardized)
abline(0,0)
abline(v=0)
```

####To conduct a cfa on the 12 item ICS (Fisher et al 2008)
###First we will create a data set with only the ICS items
```{r}
require(dplyr)
ICSDATA=select(noout, ICS1:ICS12)
head(ICSDATA)
```


###Assumptions.
####To screen for addivity. 
```{r}
correl4=cor(ICSDATA, use ="pairwise.complete.obs")
symnum(correl4)
```




####Here we will test a bi-factor model.
```{r}
ICS.CFA='
factor1=~ICS1+ICS2+ICS3+ICS5+ICS6+ICS7+ICS8+ICS12+ICS11
factor2=~ICS10+ICS4+ICS9+ICS11
Generalfactor=~1*factor1 + 1*factor2
Generalfactor~~0*factor1
Generalfactor~~0*factor2'
```

####To run the cfa for the ICS
```{r}
ICS.bifactor.fit=cfa(ICS.bifactor, data=noout,std.lv = TRUE)
lavInspect(ICS.bifactor.fit, "cov.lv")
```


####To summarize model fit
```{r}
summary(ICS.bifactor.fit, standardized=TRUE, rsquare=TRUE, fit.measure=TRUE)
```

###To produce standardized residuals for the
```{r}
zcorrel=residuals(ICS.bifactor.fit,type='standardized')
View(zcorrel$cov)
```
####To obtain model indices of the SDQ.
```{r}
modindices(ICS.bifactor.fit, sort. =TRUE, minimum.value=3.84)
```



####To reduce the bi-factor model.

```{r}
ICS.bifactor.revised='
Generalfator=~ICS1+ICS2+ICS3+ICS5+ICS6+ICS7+ICS8+ICS12+ICS11+ICS10+ICS4+ICS9
factor2=~ICS10+ICS4+ICS9+ICS11
Generalfator~~0*factor2'
```

####To run the cfa for the ICS
```{r}
ICS.bifactor.fit.revised=cfa(ICS.bifactor.revised, data=ICSDATA,std.lv = TRUE, information="observed",orthogonal=TRUE)
```


####To summarize model fit
```{r}
summary(ICS.bifactor.fit.revised, standardized=TRUE, rsquare=TRUE, fit.measure=TRUE)
```

###To produce standardized residuals for the
```{r}
zcorrel=residuals(ICS.bifactor.fit.revised,type='standardized')
View(zcorrel$cov)
```
####To obtain model indices of the SDQ.
```{r}
modindices(ICS.bifactor.fit.revised, sort. =TRUE, minimum.value=3.84)
```




