---
title: "CMVS Outcome Variables"
author: "Graham Albert"
date: "April 7 2020"
output: html_document
---

####First I will load the necessary packages in the work space.
####To load cvs files
```{r}
require(curl) 
```
######lto replace missing data
```{r}
require(mice)
```
######for advanced rotation methods
```{r}
require(GPArotation)
```
#####for data manipulation
```{r}
require(dplyr)
```
####for graphical depiction
```{r}
require(gplots)
require(gridExtra)
require(moments)
require(psych)##for EFA, and descriiptivves
require(BaylorEdPsych)####for missing value analysis
require(lavaan)####for latent variable analysis CFA and SEM.
require(semPlot)####for graphical depiction
require(BDgraph)##### for graphical depiction
require(yaml)
require(stringi)
```


####To load data set one for the CMVS.
```{r}
f <- curl("https://raw.githubusercontent.com/GrahamAlbert/Mate-Value/master/ICS_SEM_Variables.csv")
ICSSEM<- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
View(ICSSEM)
head(ICSSEM)
names(ICSSEM)
```

####To determine variable names of the data set.
```{r}
names(ICSSEM)[1]<-"ID"
names(ICSSEM)
```
####To obtain summary statistics for the data set.
```{r}
summary(ICSSEM)
```

####To obtain the descriptive statistics includinag the skewness and kurtosis values of the the cmvs
```{r}
require(psych)
normalityassessment = describe(ICSSEM[-c(1:5)])
View(normalityassessment)
```
####To generate histograms for the variables that are not normality distributed.
```{r}
hist(ICSSEM$Partner_Number)
hist(ICSSEM$Participant_Lifetime_Sex_Partners)
hist(ICSSEM$Past_Year_Sex_Partners)
hist(ICSSEM$Past_Month_Sex)
```
####To mutate the data set to conduct a log 10 transformations for the skewed outcome variables
```{r}
ICSSEM2= mutate(ICSSEM, LGPartnerNumber=log10(Partner_Number+1), LGParticipantLifetimeSexPartners = log10(Participant_Lifetime_Sex_Partners+1),LGPastYearSexPartners=log10(Past_Year_Sex_Partners+1),LGPastMonthSex = log10(Past_Month_Sex+1))
View(ICSSEM2)
```
####To generate histograms for the variables that are not normality distributed.
```{r}
hist(ICSSEM2$LGPartnerNumber)
hist(ICSSEM2$LGParticipantLifetimeSexPartners)
hist(ICSSEM2$LGPastYearSexPartners)
hist(ICSSEM2$LGPastMonthSex)
```

###To evaluate the skew of the transformed variables.
```{r}
require(psych)
describe(ICSSEM2$LGPartnerNumber)
describe(ICSSEM2$LGParticipantLifetimeSexPartners)
describe(ICSSEM2$LGPastYearSexPartners)
describe(ICSSEM2$LGPastMonthSex)
```



####To remove redundant variables
```{r}
ICSSEM3 = ICSSEM2[-c(5:7,9,12,13,14,15)]
View(ICSSEM3)
```

####To screening for missing values.
####To write a funtion to check for missng data.
```{r}
percentmiss=function(x){sum(is.na(x)/length(x))*100}
```


####To check for missing data by columns
```{r}
apply(ICSSEM3,2,percentmiss)
```
####It is less than 2% missing for all variables.

####To check for missing data by row
```{r}
missing=apply(ICSSEM3, 1,percentmiss)
table(missing)
```


####To exclude participants who are missing greater than 5% of their data.
```{r}
replacepeople=subset(ICSSEM3,missing<=5)
```
####I will be elimating three cases.

####Make sure that the columns are not missing too much data.
```{r}
apply(replacepeople, 2, percentmiss)
```

#####After excluding the cases with too much missing data we will replace the remaining missing data with a stocastic imputation method.
```{r}
require(mice)
tempnomiss = mice(replacepeople)
nomiss = complete(tempnomiss, 1)
summary(nomiss)
```
###To lookk at data within missing replaced.
```{r}
View(nomiss)
```
####To screen for multivariate outliers.
####To screen for multivariate outliers in the data set nomiss we will use mahalanobis distance statistic.
```{r}
mahal = mahalanobis(nomiss,
                    colMeans(nomiss),
                    cov(nomiss))
cutoff = qchisq(0.999, ncol(nomiss)) #####generates cutoff score
ncol(nomiss) #####determines df
table(mahal<cutoff)
```
####To exelucde the 17  outliers from the analysis.
```{r}
noout = subset(nomiss, mahal<cutoff)
View(noout)
```
####Therefore for this analysis we will have a total of 351 cases.
####T subset noout according to participants reported ethnicity.
```{r}

reduced_noout = subset(noout, Participant_Ethnicity_1 == 1 | Participant_Ethnicity_1 == 3)
View(reduced_noout)

```

####To remove the PNA from relationship status
```{r}
noout  = subset(noout, Participant_LTR !=3)
View(noout)
```


####To label the factors.
```{r}
noout$Face_Set_Code = factor(noout$Participant_Sex,
                             levels = c(1,2),
                             labels = c("Male", "Female"))
table(noout$Participant_Sex)
noout$Participant_Ethnicity_1 = factor(noout$Participant_Ethnicity_1,
                             levels = c(1,2,3,4,5,6,7,8),
                             labels = c("White", "Asian", "South Asian", "South East Asian", "West Asian", "Black African", "Native Aborginal", "Latino"))
table(noout$Participant_Ethnicity_1)
noout$Participant_LTR = factor(noout$Participant_LTR,
                             levels = c(1,2),
                             labels = c("Yes", "No"))
table(noout$Participant_LTR)
noout$Had_Sex = factor(noout$Had_Sex,
                             levels = c(1,2),
                             labels = c("Yes", "No"))
table(noout$Had_Sex)
```


###Assumptions.
####To screen for multcorinately
####Althoughy the items are correlated (as they should) we do not have multi-colinearity so we proceed with SEM assumptionn set.

```{r}
correl=cor(noout[c(8:79)], use ="pairwise.complete.obs")
View(correl)
```
###EFA assumption set up.
```{r}
random = rchisq(nrow(noout[c(8:79)]),7)
fake = lm(random ~.,data=noout[c(8:79)])
standardized = rstudent(fake)
fitted = scale(fake$fitted.values)
```
####To test for normality.
```{r}
hist(standardized)
```
####The data is normal but slightly skewed to the right

####It appears as though we have multi-variate normality so we proceed to test linearity.
####To test for linearity.
```{r}
qqnorm(standardized)
abline(0,1)
```
####Our variables are linear so we procede to test homogenity of vairance.


####Homogenity of Variance.
```{r}
plot(fitted, standardized)
abline(0,0)
abline(v=0)
```





####CFA
###Three factor
###We will revise the model to include correlated error varainces between items. 

```{r}
three.model='
PARTUP=~MEQ1+MEQ3+MEQ5+MEQ6+MEQ7+MEQ10
MEQ7~~MEQ10
MEQ3~~MEQ5
MEQ7~~MEQ5
SINGLE=~MEQ8+MEQ6+MEQ2+MEQ12
MATE=~MEQ11+MEQ9+MEQ4'
```


####To run the three factor moddel with correlated residuals.
```{r}
require(lavaan)
three.fit=cfa(three.model, data=noout)
```

###To check for heywood and to obain summary statistics for the MEQ model
```{r}
require(lavaan)
summary(three.fit, standardized=TRUE, rsquare=TRUE, fit.measure=TRUE)
```

###standardized residuals three
```{r}
zcorrel=residuals(three.fit,type='standardized')
View(zcorrel$cov)
```
####to obtain model indices.
```{r}
modindices(three.fit, sort. =TRUE, minimum.value=3.84)
```
####To obtian standarized and unstandardized factor laodings
```{r}
parameterEstimates(three.fit)
standardizedsolution(three.fit)
```


####To run a three factor model for the SOI-R

####To specify the model.
```{r}
SOI.three.model='
behavior=~SOIR2+SOIR3+SOIR1
attitude=~SOIR4+SOIR5+SOIR6
desire=~SOIR8+SOIR9+SOIR7'
```
####To run the cfa for the SOIR
```{r}
require(lavaan)
SOI.three.fit=cfa(SOI.three.model, data=noout)
```


####To summarize model fit of the SOIR
```{r}
summary(SOI.three.fit, standardized=TRUE, rsquare=TRUE, fit.measure=TRUE)
```
###standardized residuals for the SOIR
```{r}
zcorrel=residuals(SOI.three.fit,type='standardized')
View(zcorrel$cov)
```
####To obtain model indices.
```{r}
modindices(SOI.three.fit, sort. =TRUE, minimum.value=3.84)
```

####To obtian standarized and unstandardized factor laodings
```{r}
parameterEstimates(SOI.three.fit)
standardizedsolution(SOI.three.fit)
```


###Based on the inspection of the items it seems as though there are two factors factor one deals with a fear of being worse than intrasexual competitors (items:1,2,3,5,6,7,8,12,11) and wanting to be better than competitor (items:4,9,10,11)

####To specify the two factor model.
```{r}
ICS.two.model='
FEARWORSE=~ICS1+ICS2+ICS3+ICS5+ICS6+ICS7+ICS8+ICS12+ICS11
BEBETTER=~ICS10+ICS4+ICS9+ICS11'
```

####To run the cfa for the ICS
```{r}
require(lavaan)
ICS.two.fit=cfa(ICS.two.model, data=noout)
```


####To summarize model fit
```{r}
summary(ICS.two.fit, standardized=TRUE, rsquare=TRUE, fit.measure=TRUE)
```


###To produce standardized residuals for the
```{r}
zcorrel=residuals(ICS.two.fit,type='standardized')
View(zcorrel$cov)
```
####To obtain model indices of the ICS
```{r}
modindices(ICS.two.fit, sort. =TRUE, minimum.value=3.84)
```

####To obtian standarized and unstandardized factor loadings
```{r}
parameterEstimates(ICS.two.fit, standardized =T)
standardizedsolution(ICS.two.fit)
```








#To conduct confirmator factor analysis for the MRI-SF.
```{r}
MATE_RETENTION_MODEL= '
COSTINFLICT=~MRSF38+MRSF36+MRSF2+MRSF4+MRSF8+MRSF17+MRSF19+MRSF21+MRSF23+MRSF27+MRSF33
BENEFITPROV=~MRSF30+MRSF9+MRSF11+MRSF12+MRSF14+MRSF15+MRSF28+MRSF29+MRSF31+MRSF33
MATEGAURD=~MRSF6+MRSF3+MRSF1+MRSF7+MRSF10+MRSF22+MRSF25+MRSF14
VIGILANCE=~MRSF5+MRSF18+MRSF24+MRSF37+MRSF11
POSSESSION=~MRSF16+MRSF26+MRSF35+MRSF9
MRSF38~~MRSF19'
```



####run the models.
```{r}
require(lavaan)
MATE_RETENTION_FIT=cfa(MATE_RETENTION_MODEL, data=noout)
```

###to obtain summary statistics   model.
```{r}
require(lavaan)
summary(MATE_RETENTION_FIT, standardized=TRUE, rsquare=TRUE, fit.measure=TRUE)
```

###standardized residuals three
```{r}
zcorrel=residuals(MATE_RETENTION_FIT,type='standardized')
View(zcorrel$cov)
```
####o obtain model indices.
```{r}
modindices(MATE_RETENTION_FIT, sort. =TRUE, minimum.value=3.84)
```

####To obtian standarized and unstandardized factor laodings
```{r}
parameterEstimates(MATE_RETENTION_FIT)
standardizedsolution(MATE_RETENTION_FIT)
```


####################

##To conduct SEMs in which we regress age, relationship status and ethnicity onto the factors of the ICS.
###Here we are testing to see if age predicts 

####################

#Analysis controlling for age.


####To specify the two factor model with age as a predictor.
```{r}
ICS.AGE='
FEARWORSE=~ICS1+ICS2+ICS3+ICS5+ICS6+ICS7+ICS8+ICS12+ICS11
BEBETTER=~ICS10+ICS4+ICS9+ICS11
FEARWORSE~Participant_Age
BEBETTER~Participant_Age'
```

####To run the cfa for the ICS
```{r}
require(lavaan)
ICS.AGE.FIT=sem(ICS.AGE, data=noout)
```


####To summarize model fit
```{r}
summary(ICS.AGE.FIT, standardized=TRUE, rsquare=TRUE, fit.measure=TRUE)
```


###To produce standardized residuals for the
```{r}
zcorrel=residuals(ICS.AGE.FIT,type='standardized')
View(zcorrel$cov)
```
####To obtain model indices of the ICS
```{r}
modindices(ICS.AGE.FIT, sort. =TRUE, minimum.value=3.84)
```

####To obtian standarized and unstandardized factor loadings
```{r}
parameterEstimates(ICS.AGE.FIT, standardized =T)
standardizedsolution(ICS.AGE.FIT)
```


#To control for relationship status.
#To subset the data to only include those who report either being in a relationship or not.
```{r}
ICSAGERELDATA = subset(noout,Participant_LTR!='PNA')
View(ICSAGERELDATA)
```
###To factor LTR
```{r}
ICSAGERELDATA$Participant_LTR = as.factor(ICSAGERELDATA$Participant_LTR)
nlevels(ICSAGERELDATA$Participant_LTR)

```
####To specify the two factor model with relatioship status
```{r}
ICS.LTR='
FEARWORSE=~ICS1+ICS2+ICS3+ICS5+ICS6+ICS7+ICS8+ICS12+ICS11
BEBETTER=~ICS10+ICS4+ICS9+ICS11
FEARWORSE~Participant_LTR
BEBETTER~Participant_LTR'
```

####To run the cfa for the ICS
```{r}
require(lavaan)
ICS.LTR.FIT=sem(ICS.LTR, data=ICSAGERELDATA)
```


####To summarize model fit
```{r}
summary(ICS.LTR.FIT, standardized=TRUE, rsquare=TRUE, fit.measure=TRUE)
```


###To produce standardized residuals for the
```{r}
zcorrel=residuals(ICS.LTR.FIT,type='standardized')
View(zcorrel$cov)
```
####To obtain model indices of the ICS
```{r}
modindices(ICS.LTR.FIT, sort. =TRUE, minimum.value=3.84)
```

####To obtian standarized and unstandardized factor loadings
```{r}
parameterEstimates(ICS.LTR.FIT, standardized =T)
standardizedsolution(ICS.LTR.FIT)
```












####T subset noout according to participants reported ethnicity.
```{r}
reduced_noout$Participant_Ethnicity_1 = factor(reduced_noout$Participant_Ethnicity_1,
                             levels = c(1,3),
                             labels = c("White","South  Asian"))
table(reduced_noout$Participant_Ethnicity_1)
```

####To specify the two factor model.
#####Here we regress partipants ethnicity onto the two factors.
```{r}
ICS.ETH='
FEARWORSE=~ICS1+ICS2+ICS3+ICS5+ICS6+ICS7+ICS8+ICS12+ICS11
BEBETTER=~ICS10+ICS4+ICS9+ICS11
FEARWORSE~Participant_Ethnicity_1
BEBETTER~Participant_Ethnicity_1'
```

####To run the cfa for the ICS
```{r}
require(lavaan)
ICS.ETH.FIT=sem(ICS.ETH, data=reduced_noout)
```


####To summarize model fit
```{r}
summary(ICS.ETH.FIT, standardized=TRUE, rsquare=TRUE, fit.measure=TRUE)
```


###To produce standardized residuals for the
```{r}
zcorrel=residuals(ICS.ETH.FIT,type='standardized')
View(zcorrel$cov)
```
####To obtain model indices of the ICS
```{r}
modindices(ICS.ETH.FIT, sort. =TRUE, minimum.value=3.84)
```

####To obtian standarized and unstandardized factor loadings
```{r}
parameterEstimates(ICS.ETH.FIT, standardized =T)
standardizedsolution(ICS.ETH.FIT)
```


#To conduct an SEM including the factors from the MRI-SF, the SOI-R, and the MEQ.
###The factors from the SOI-R attitude and desire will be specified as predicting the factors from the ICS.
###The factors of the ICS will  be specified as predicting the factors of the MEQ and MRI-SF
```{r}
SEMMODEL='
COSTINFLICT=~MRSF38+MRSF36+MRSF2+MRSF4+MRSF8+MRSF17+MRSF19+MRSF21+MRSF23+MRSF27+MRSF33
BENEFITPROV=~MRSF30+MRSF9+MRSF11+MRSF12+MRSF14+MRSF15+MRSF28+MRSF29+MRSF31+MRSF33
COSTINFLICT~~BENEFITPROV
PARTUP=~MEQ1+MEQ3+MEQ5+MEQ6+MEQ7+MEQ10
SINGLE=~MEQ8+MEQ6+MEQ2+MEQ12
PARTUP~~SINGLE
COSTINFLICT~~PARTUP
COSTINFLICT~~SINGLE
BENEFITPROV~~SINGLE
BENEFITPROV~~PARTUP
FEARWORSE=~ICS1+ICS2+ICS3+ICS5+ICS6+ICS7+ICS8+ICS12+ICS11
BEBETTER=~ICS10+ICS4+ICS9+ICS11
FEARWORSE~~BEBETTER
attitude=~SOIR4+SOIR5+SOIR6
desire=~SOIR8+SOIR9+SOIR7
attitude~~desire
FEARWORSE~attitude+desire
BEBETTER~attitude+desire
COSTINFLICT~FEARWORSE+attitude+desire
BENEFITPROV~BEBETTER+attitude+desire
PARTUP~FEARWORSE+BEBETTER+attitude+desire
SINGLE~FEARWORSE+BEBETTER+attitude+desire'
```


####To run the cfa for the ICS
```{r}
require(lavaan)
SEM.FIT=sem(SEMMODEL, data=noout)
```


####To summarize model fit
```{r}
summary(SEM.FIT, standardized=TRUE, rsquare=TRUE, fit.measure=TRUE)
```


###To produce standardized residuals for the
```{r}
zcorrel=residuals(SEM.FIT,type='standardized')
View(zcorrel$cov)
```
####To obtain model indices of the ICS
```{r}
modindices(SEM.FIT, sort. =TRUE, minimum.value=3.84)
```

####To obtian standarized and unstandardized factor laodings
```{r}
parameterEstimates(SEM.FIT)
standardizedsolution(SEM.FIT)
```



























#####To create a model in which the ICS factors predict indivdiuals sexual behavior.
```{r}
SEM.Model.Reproductive='
FEARWORSE=~ICS1+ICS2+ICS3+ICS5+ICS6+ICS7+ICS8+ICS12+ICS11
BEBETTER=~ICS10+ICS4+ICS9+ICS11
BEBETTER~~FEARWORSE
Age_Sex~FEARWORSE+BEBETTER+Participant_Sex
LGPartnerNumber~FEARWORSE+BEBETTER+Participant_Age
LGParticipantLifetimeSexPartners~FEARWORSE+BEBETTER+Participant_Age
LGPastYearSexPartners~FEARWORSE+BEBETTER+Participant_LTR
LGPastMonthSex~FEARWORSE+BEBETTER+Participant_LTR
Participant_LTR~FEARWORSE+BEBETTER'
```
###To edit noout to 
```{r}
require(dplyr)
noout = mutate(noout, Participant_LTR = as.numeric(Participant_LTR))
View(noout)
```



####To run the cfa for the ICS
```{r}
require(lavaan)
SEM.fit.Reproductive=sem(SEM.Model.Reproductive, data=noout)
```

####to depict the  models.
```{r}
require(BGgraphs)
require(semPlot)
semPaths(SEM.fit.Reproductive, whatLabels="std",layout="tree",edge.label.cex = 0.5,curvePivot = TRUE)
```

####To summarize model fit
```{r}
summary(SEM.fit.Reproductive, standardized=TRUE, rsquare=TRUE, fit.measure=TRUE)
```

###To produce standardized residuals for the
```{r}
zcorrel=residuals(SEM.fit.Reproductive,type='standardized')
View(zcorrel$cov)
```
####To obtain model indices of the SDQ.
```{r}
modindices(SEM.fit.Reproductive, sort. =TRUE, minimum.value=3.84)
```

####To obtian standarized and unstandardized factor loadings
```{r}
parameterEstimates(SEM.fit.Reproductive, standardized =T)
standardizedsolution(SEM.fit.Reproductive)
```

