---
title: "Intrasexually Competitie Attitude Scale EFA"
author: "Graham Albert"
date: "January 8, 2019"
output: html_document
---

####First I will load the necessary packages in the work space.
####To load cvs files
```{r}
require(curl) 
```
######lto replace missing data
```{r}
require(mice)
```
######for advanced rotation methods
```{r}
require(GPArotation)
```
#####for data manipulation
```{r}
require(dplyr)
```
####for graphical depiction
```{r}
require(gplots)
require(gridExtra)
require(moments)
require(psych)##for EFA, and descriiptivves
require(BaylorEdPsych)####for missing value analysis
require(lavaan)####for latent variable analysis CFA and SEM.
require(semPlot)####for graphical depiction
require(BDgraph)##### for graphical depiction
require(yaml)
require(stringi)
```


####To load data set one for the CMVS.
```{r}
f <- curl("https://raw.githubusercontent.com/GrahamAlbert/Mate-Value/master/Intrasexually_comeptitive_attitude_EFA_20190108.csv")
ICSEFA<- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
View(ICSEFA)
head(ICSEFA)
names(ICSEFA)
```

####To determine variable names of the data set.
```{r}
names(ICSEFA)[1]<-"SEX"
names(ICSEFA)
```
####To obtain summary statistics for the data set.
```{r}
summary(ICSEFA)
```
####To remove categorical variables.
```{r}
require(dplyr)
ICSEFA2 = ICSEFA[-c(3:13)]
View(ICSEFA2)
```
####To obtain the descriptive statistics includinag the skewness and kurtosis values of the the ICS
```{r}
require(psych)
normalityassessment = describe(ICSEFA2[-c(1,2)])
View(normalityassessment)
```
####All items are normally distributed.
####We will proceed to analyze patterns of missingness.







####To screening for missing values.
####To write a funtion to check for missng data.
```{r}
percentmiss=function(x){sum(is.na(x)/length(x))*100}
```


####To check for missing data by columns
```{r}
apply(ICSEFA2,2,percentmiss)
```
####It is less than 2% missing for all variables.

####To check for missing data by row
```{r}
missing=apply(ICSEFA2, 1,percentmiss)
table(missing)
```


####To exclude participants who are missing greater than 5% of their data.
```{r}
replacepeople=subset(ICSEFA2,missing<=5)
```
####I will be elimating three cases.




####Make sure that the columns are not missing too much data.
```{r}
apply(replacepeople, 2, percentmiss)
```

###After inspecting the skewness and kurtosis statistics for the cmvs it appears that all questions are relatively normally distributed therefore I will proceed with analyzing the data for missing values.
####To analysis the data set for missing value patters
```{r}
require(BaylorEdPsych)
require(mvnmle)
LittleMCAR(replacepeople[-c(1,2)])
```
#####After excluding the cases with too much missing data we will replace the remaining missing data with a stocastic imputation method.
```{r}
require(mice)
tempnomiss = mice(replacepeople)
nomiss=complete (tempnomiss, 1)
summary(nomiss)
```
###To lookk at data within missing replaced.
```{r}
View(nomiss)
```
####To screen for multivariate outliers.
####To screen for multivariate outliers in the data set nomiss we will use mahalanobis distance statistic.
```{r}
mahal = mahalanobis(nomiss,
                    colMeans(nomiss),
                    cov(nomiss))
cutoff = qchisq(0.999, ncol(nomiss)) #####generates cutoff score
ncol(nomiss) #####determines df
table(mahal<cutoff)
```
####To exelucde the 18  outliers from the analysis.
```{r}
noout = subset(nomiss, mahal<cutoff)
View(noout)
```
####Therefore for this analysis we will have a total of 263 cases.


###Assumptions.
####To screen for multcorinately
####Althoughy the items are correlated (as they should) we do not have multi-colinearity so we proceed with EFA assumptionn set.
####I will be excluding the single report items from rows 1 and 2(sex and age) from analysis, so that I can look at the correlations amongst the items.
```{r}
correl=cor(noout[-c(1,2)], use ="pairwise.complete.obs")
View(correl)
```
###EFA assumption set up.
```{r}
random = rchisq(nrow(noout[-c(1,2)]),7)
fake = lm(random ~.,data=noout[-c(1,2)])
standardized = rstudent(fake)
fitted = scale(fake$fitted.values)
```
####To test for normality.
```{r}
hist(standardized)
```
####The data is normal but slightly skewed to the right

####It appears as though we have multi-variate normality so we proceed to test linearity.
####To test for linearity.
```{r}
qqnorm(standardized)
abline(0,1)
```
####Our variables are linear so we procede to test homogenity of vairance.


####Homogenity of Variance.
```{r}
plot(fitted, standardized)
abline(0,0)
abline(v=0)
```


####To conduct a cfa on the 12 item ICS (Fisher et al 2008)
###First we will create a data set with only the ICS items
```{r}
detach("package:BDgraph", unload=TRUE)
ICSDATA=select(noout, ICS1:ICS12)
head(ICSDATA)
```


###Assumptions.
####To screen for addivity for the ICS
```{r}
correl4=cor(ICSDATA, use ="pairwise.complete.obs")
symnum(correl4)
```
###To run bartletts test of sphericty
```{r}
require(psych)
cortest.bartlett(correl4, n = nrow(ICSDATA))
```
####sampling adequacy KMO test.
```{r}
KMO(correl4)
```
####detereminig number of factors for extraction.
```{r}
nofactors <- fa.parallel(ICSDATA,fm="ml",fa="fa")
sum(nofactors$fa.values > 1.0)#####old kaiser criterion
sum(nofactors$fa.values > 0.7)#####new kaiser criterion
```
####Kaisers old criterion and the screen plot recommends 1 factor for the ICS,whereas Kaisers new criterion and the parrallel analysis recommends 2 factors.
####We will start with a 1 factor solution.

```{r}
require(GPArotation)
ICSefa1 = fa(ICSDATA,nfactors=1,rotate="oblimin", fm="ml")
ICSefa1
```

####To compute the CFI of the ICS.
```{r}
CFI<-1-((ICSefa1$STATISTIC-ICSefa1$dof)/(ICSefa1$null.chisq-ICSefa1$null.dof))
CFI
```

####The two factor solution for the ICS
```{r}
require(GPArotation)
ICSefafinal = fa(ICSDATA,nfactors=2,rotate="oblimin", fm="ml")
ICSefafinal
```



####To compute the CFI of the SDQ.
```{r}
CFI<-1-((ICSefafinal$STATISTIC-ICSefafinal$dof)/(ICSefafinal$null.chisq-ICSefafinal$null.dof))
CFI
```
####To create a factor plot.
```{r}
factor.plot(ICSefafinal, labels=rownames(ICSefafinal)) 
```   


#To create a diagram of the analysis.
```{r}
fa.diagram(ICSefafinal, simple=FALSE)
```



####Now that we have explored the factor structure we will load the second round of data and conduct a CFA and an MGFA testing measurement invariance and population heterogenity.


